---
output: rmarkdown::github_document
bibliography: README-refs.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`shifttx`

[![Travis-CI Build Status](https://travis-ci.org/nhejazi/shifttx.svg?branch=master)](https://travis-ci.org/nhejazi/shifttx)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/nhejazi/shifttx?branch=master&svg=true)](https://ci.appveyor.com/project/nhejazi/shifttx)
[![Coverage Status](https://img.shields.io/codecov/c/github/nhejazi/shifttx/master.svg)](https://codecov.io/github/nhejazi/shifttx?branch=master)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)

> Targeted Learning of Continuous Interventions with Stochastic Treatment
> Regimes

__Authors:__ [Nima Hejazi](http://nimahejazi.org) and [David
Benkeser](https://www.benkeserstatistics.com/)

---

## What's `shifttx`?

`shifttx` is an R package that makes it easy to compute targeted minimum
loss-based estimates of the population-level causal effects of interventions
based on stochastic mechanisms for treatment assignment.

* Original estimator and iterative algorithm proposed in @munoz2012population
* One-step estimation procedure and algorithm introduced in @diaz2017stochastic.
* See @vdl2017targeted for a discussion on recent developments in Targeted
  Learning.
* See @vdl2011targeted for an introduction to Targeted Learning.

---

## Installation

Install the most recent _stable release_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-master-installation, eval = FALSE}
devtools::install_github("nhejazi/shifttx")
```

---

## Example

To illustrate how `shifttx` may be used to ascertain the effect of a treatment,
consider the following example:

```{r example}
library(shifttx)

# first, define simulation parameters and observed data
n <- 100
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.7))
A <- rpois(n, lambda = exp(3 + 0.3 * log(W$W1) - 0.2 * exp(W$W1) * W$W2))
Y <- rbinom(n, 1, plogis(-1 + 0.05 * A - 0.02 * A * W$W2 + 0.2 * A * tan(W$W1^2)
                         - 0.02 * W$W1 * W$W2 + 0.1 * A * W$W1 * W$W2))
fitA.0 <- glm(A ~ I(log(W1)) + I(exp(W1)):W2, family = poisson,
              data = data.frame(A, W))
fitY.0 <- glm(Y ~ A + A:W2 + A:I(tan(W1^2)) + W1:W2 + A:W1:W2,
              family = binomial, data = data.frame(A, W))

# next, we define the treatment mechanisma and outcome in the true model
gn.0  <- function(A = A, W = W) {
  dpois(A, lambda = predict(fitA.0, newdata = W, type = "response"))
}
Qn.0 <- function(A = A, W = W) {
  predict(fitY.0, newdata = data.frame(A, W, row.names = NULL),
          type = "response")
}

# finally, we may compute targeted estimates of the intervention effect
est_target <- tmle_shift(Y = Y, A = A, W = W, Qn = Qn.0, gn = gn.0, delta = 2,
                         tol = 1e-4, iter_max = 5, A_val = seq(1, 60, 1))

# let's just examine the first few results
head(est_target)
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/nhejazi/shifttx/issues).

---

## License

&copy; 2017 [Nima S. Hejazi](http://nimahejazi.org)

The contents of this repository are distributed under the MIT license. See below
for details:
```
MIT License

Copyright (c) 2017 Nima S. Hejazi

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```

---

## References

