---
output:
  rmarkdown::github_document
bibliography: "README-refs.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# R/`txshift`

[![Travis-CI Build Status](https://travis-ci.org/nhejazi/txshift.svg?branch=master)](https://travis-ci.org/nhejazi/txshift)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/nhejazi/txshift?branch=master&svg=true)](https://ci.appveyor.com/project/nhejazi/txshift)
[![Coverage Status](https://img.shields.io/codecov/c/github/nhejazi/txshift/master.svg)](https://codecov.io/github/nhejazi/txshift?branch=master)
[![CRAN](http://www.r-pkg.org/badges/version/txshift)](http://www.r-pkg.org/pkg/txshift)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/txshift)](https://CRAN.R-project.org/package=txshift)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)

> Targeted Learning of the Causal Effects of Stochastic Interventions

__Authors:__ [Nima Hejazi](https://nimahejazi.org) and [David
Benkeser](https://www.benkeserstatistics.com/)

---

## What's `txshift`?

The `txshift` R package is designed to compute targeted maximum likelihood (TML)
estimates of the counterfactual mean of an outcome under stochastic mechanisms
for treatment assignment and related causal parameters (@diaz2012population). In
particular, `txshift ` implements and builds upon a simplified algorithm for the
TML estimator proposed by @diaz2018stochastic.

For many practical applications (e.g., vaccine efficacy trials), it is often the
case that the observed data structure is generated under a two-phase sampling
mechanism (i.e., a two-stage design). In such cases, TML estimators must be
augmented to exhibit efficiency in spite of the challenges induced by the
censoring process. An appropriate augmentation procedure was first proposed by
@rose2011targeted2sd, who proposed the use of inverse probability of censoring
weights (IPCW) alongside an augmentation of the relevant efficient influence
function. `txshift` extends this approach to computing one-step and IPCW-TML
estimators of the counterfactual mean under a stochastic treatment regime.

---

## Installation

Install the most recent _stable release_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-master-installation, eval = FALSE}
devtools::install_github("nhejazi/txshift", build_vignettes = FALSE)
```

---

## Example

To illustrate how `txshift` may be used to ascertain the effect of a treatment,
consider the following example:

```{r example, message=FALSE, warning=FALSE}
library(txshift)
library(condensier)
set.seed(429153)

# simulate simple data for tmle-shift sketch
n_obs <- 1000  # number of observations
n_w <- 1  # number of baseline covariates
p_w <- 0.5  # probability of a success ("1") in the baseline variables
tx_mult <- 2  # multiplier for the effect of W = 1 on the treatment

## baseline covariate -- simple, binary
W <- as.numeric(replicate(n_w, rbinom(n_obs, 1, p_w)))

## create treatment based on baseline W
A <- as.numeric(rnorm(n_obs, mean = tx_mult * W, sd = 1))

# create outcome as a linear function of A, W + white noise
Y <- A + W + rnorm(n_obs, mean = 0, sd = 1)

# fit the TMLE
tmle_shift <- tmle_txshift(W = W, A = A, Y = Y, delta = 0.5,
                           g_fit_args = list(fit_type = "glm",
                                             nbins = 25,
                                             bin_method = "dhist",
                                             bin_estimator =
                                               speedglmR6$new(),
                                             parfit = FALSE),
                           Q_fit_args = list(fit_type = "glm",
                                             glm_formula = "Y ~ .")
                          )

# conveniently summarize the results
summary(tmle_shift)

# now, let's introduce a censoring process (for two-stage sampling)
C <- rbinom(n_obs, 1, plogis(W + Y))

# fit an IPCW-TMLE to account for this censoring process:
ipcwtmle_shift <- tmle_txshift(W = W, A = A, Y = Y, delta = 0.5,
                               C = C, V = c("W", "Y"),
                               max_iter = 10,  # limit iterations for speed
                               ipcw_fit_args = list(fit_type = "glm"),
                               g_fit_args = list(fit_type = "glm",
                                                 nbins = 25,
                                                 bin_method = "dhist",
                                                 bin_estimator =
                                                   speedglmR6$new(),
                                                 parfit = FALSE),
                               Q_fit_args = list(fit_type = "glm",
                                                 glm_formula = "Y ~ ."),
                               eif_reg_spec = FALSE  # fit EIF with a GLM
                              )

# conveniently summarize the results
summary(ipcwtmle_shift)
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/nhejazi/txshift/issues).

---

## Contributions

Contributions are very welcome. Interested contributors should consult our
[contribution
guidelines](https://github.com/nhejazi/txshift/blob/master/CONTRIBUTING.md)
prior to submitting a pull request.

---

## Citation

After using the `txshift` R package, please cite the following:

        @manual{hejazi2019txshift,
          author = {Hejazi, Nima S and Benkeser, David C},
          title = {txshift: {Targeted Learning} of the Causal Effects of
            Stochastic Interventions in {R}},
          year  = {2019},
          url = {https://github.com/nhejazi/txshift},
          note = {R package version 0.2.2}
        }

---

## Related

* [R/`tmle3shift`](https://github.com/tlverse/tmle3shift) - An R package
  providing an independent implementation of the same core routines for the TML
  estimation procedure and statistical methodology as is made available here,
  through reliance on a unified interface for Targeted Learning provided by the
  [`tmle3`](https://github.com/tlverse/tmle3) engine of the [`tlverse`
  ecosystem](https://github.com/tlverse).

* [R/`medshift`](https://github.com/nhejazi/medshift) - An R package providing
  facilities to estimate the causal effect of stochastic treatment regimes in
  the mediation setting, including classical (IPW) and augmented double robust
  (one-step) estimators. This is an implementation of the methodology explored
  in @diaz2019causal.

* [R/`condensier`](https://github.com/osofr/condensier) - Estimation of the
  treatment mechanism component of this parameter requires conditional density
  estimation, which is implemented rather generally (for a variety of hazard
  regression strategies and regression functions) in this package. This package
  implements a variant of the methodology proposed in @diaz2011super.

* [R/`haldensify`](https://github.com/nhejazi/haldensify) - A minimal package
  for estimating the conditional density treatment mechanism component of this
  parameter based on using the [highly adaptive
  lasso](https://github.com/tlverse/hal9001) for the required pooled hazard
  regression. This package implements the methdology proposed in @diaz2011super.

---

## Funding

The development of this software was supported in part through a grant from the
National Institutes of Health: [T32
LM012417-02](https://projectreporter.nih.gov/project_info_description.cfm?aid=9248418&icde=37849831&ddparam=&ddvalue=&ddsub=&cr=1&csb=default&cs=ASC&pball=).

---

## License

&copy; 2017-2019 [Nima S. Hejazi](https://nimahejazi.org)

The contents of this repository are distributed under the MIT license. See below
for details:
```
MIT License

Copyright (c) 2017-2019 Nima S. Hejazi

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```

---

## References

