---
output:
  rmarkdown::github_document
bibliography: "`r here::here('README-refs.bib')`"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`txshift`

[![Travis-CI Build Status](https://travis-ci.org/nhejazi/txshift.svg?branch=master)](https://travis-ci.org/nhejazi/txshift)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/nhejazi/txshift?branch=master&svg=true)](https://ci.appveyor.com/project/nhejazi/txshift)
[![Coverage Status](https://img.shields.io/codecov/c/github/nhejazi/txshift/master.svg)](https://codecov.io/github/nhejazi/txshift?branch=master)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)

> Targeted Learning of Continuous Intervention Effects with Stochastic Treatment
> Regimes

__Authors:__ [Nima Hejazi](https://nimahejazi.org) and [David
Benkeser](https://www.benkeserstatistics.com/)

---

## What's `txshift`?

The `txshift` R package is designed to provide facilities to compute targeted
maximum likelihood estimates (TMLE) of the population-level causal effect of
interventions based on stochastic mechanisms for treatment assignment
(@munoz2012population). As opposed to the original algorithm given for computing
such a TMLE, `txshift ` implements and builds upon subsequent work by
@diaz2018stochastic, who reveal a simplified algorithm for computing the TML
estimator of the shift intervention causal effect parameter.

For many practical applications (e.g., observational medical studies), it is
often the case that the data structure of interest is generated by a selective
sampling process (i.e., a two-stage design). In such cases, TML estimators may
be augmented to cope with -- and even exhibit efficiency -- in spite of the
challenges induced by such an artificial censoring process. This augmentation
procedure is done by way of adding inverse probability of censoring weights
(IPCW), which leads to an IPCW-TMLE, originally proposed by
@rose2011targeted2sd. `txshift` extends the approach of computing IPCW-TMLEs to
the shift intervention causal effect parameter.

For background on the Targeted Learning methodology, as well as recent advances,
the canonical references are @vdl2011targeted and @vdl2018targeted.

---

## Installation

Install the most recent _stable release_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-master-installation, eval = FALSE}
devtools::install_github("nhejazi/txshift", build_vignettes = FALSE)
```

---

## Example

To illustrate how `txshift` may be used to ascertain the effect of a treatment,
consider the following example:

```{r example, message=FALSE, warning=FALSE}
library(txshift)
library(condensier)
set.seed(429153)

# simulate simple data for tmle-shift sketch
n_obs <- 1000  # number of observations
n_w <- 1  # number of baseline covariates
p_w <- 0.5  # probability of a success ("1") in the baseline variables
tx_mult <- 2  # multiplier for the effect of W = 1 on the treatment

## baseline covariate -- simple, binary
W <- as.numeric(replicate(n_w, rbinom(n_obs, 1, p_w)))

## create treatment based on baseline W
A <- as.numeric(rnorm(n_obs, mean = tx_mult * W, sd = 1))

# create outcome as a linear function of A, W + white noise
Y <- A + W + rnorm(n_obs, mean = 0, sd = 1)

# fit the TMLE
tmle_shift <- tmle_txshift(W = W, A = A, Y = Y, delta = 0.5,
                           g_fit_args = list(fit_type = "glm",
                                             nbins = 25,
                                             bin_method = "dhist",
                                             bin_estimator =
                                               speedglmR6$new(),
                                             parfit = FALSE),
                           Q_fit_args = list(fit_type = "glm",
                                             glm_formula = "Y ~ .")
                          )

# conveniently summarize the results
summary(tmle_shift)

# now, let's introduce a censoring process (for two-stage sampling)
C <- rbinom(n_obs, 1, plogis(W + Y))

# fit an IPCW-TMLE to account for this censoring process:
ipcwtmle_shift <- tmle_txshift(W = W, A = A, Y = Y, delta = 0.5,
                               C = C, V = c("W", "Y"),
                               ipcw_fit_args = list(fit_type = "glm"),
                               g_fit_args = list(fit_type = "glm",
                                                 nbins = 25,
                                                 bin_method = "dhist",
                                                 bin_estimator =
                                                   speedglmR6$new(),
                                                 parfit = FALSE),
                               Q_fit_args = list(fit_type = "glm",
                                                 glm_formula = "Y ~ .")
                              )

# conveniently summarize the results
summary(ipcwtmle_shift)
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/nhejazi/txshift/issues).

---

## Contributions

Contributions are very welcome. Interested contributors should consult our
[contribution
guidelines](https://github.com/nhejazi/txshift/blob/master/CONTRIBUTING.md)
prior to submitting a pull request.

---

## Citation

After using the `txshift` R package, please cite the following:

        @manual{hejazi2018txshift,
          author = {Hejazi, Nima S and {van der Laan}, Mark J and Benkeser,
            David C},
          title = {txshift: {Targeted Learning} of stochastic intervention
            effects with robust inference in {R}},
          year  = {2018},
          url = {https://github.com/nhejazi/txshift},
          note = {R package version 0.2.0}
        }

---

## Funding

The development of this software was supported in part through a grant from the
National Institutes of Health: [T32
LM012417-02](https://projectreporter.nih.gov/project_info_description.cfm?aid=9248418&icde=37849831&ddparam=&ddvalue=&ddsub=&cr=1&csb=default&cs=ASC&pball=).

---

## License

&copy; 2017-2018 [Nima S. Hejazi](https://nimahejazi.org)

The contents of this repository are distributed under the MIT license. See below
for details:
```
MIT License

Copyright (c) 2017-2018 Nima S. Hejazi

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```

---

## References

