---
title: "IPCW-TMLEs with Stochastic Treatment Regimes"
author: "[Nima Hejazi](https://nimahejazi.org) and [David
  Benkeser](https://www.benkeserstatistics.com/)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: vignette-refs.bib
vignette: >
  %\VignetteIndexEntry{IPCW-TMLEs with Stochastic Treatment Regimes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
options(scipen = 999)
```

## Introduction

For a more general introduction to the targeted maximum likelihood estimator
introduced in @diaz2018stochastic and the corresponding software implementation
found in this package, a better resource to consult is the introductory
vignette. This document builds on that work, describing an inverse probability
of censoring weighted TML estimator (IPCW-TMLE) for the same target parameter
when a censoring process is introduced into the observed data structure. In this
case the observed data structure may be denoted $O = (W, \Delta A, Y)$, where
$W$ is a set of baseline covariates, $A$ is a continuous-valued treatment or
intervention, $Y$ is an outcome of interest, and $\Delta$ is an indicator of
missingness ($1$ corresponding to being observed, $0$ to being missing). Perhaps
the best way to appreciate the utilities provided for computing IPCW-TMLEs in
this R package is to work through examples, so let's simply jump to that.

To start, let's load the packages we'll use and set a seed for simulation:

```{r setup}
library(tidyverse)
library(data.table)
library(haldensify)
library(sl3)
library(txshift)
set.seed(429153)
```

---

## Data and Notation

1. Start with a simple additive shift -- i.e., $d(a,w) = a + \delta$ if
   $a \leq u(w) - \delta$ or $d(a, w) = a$ if $a \geq u(w) - \delta$.

2. The additive shift will have support everywhere -- i.e.,
   $a \leq u(w)$ is true everywhere.

3. Consider the case in which some observations are missing --- for now, let us
   consider a simple case-control sampling case wherein the treatment is
   partially censored. Let us denote the sampling/censoring mechanism $\Delta$,
   and let $1$ correspond to being observed and $0$ to being missing.

4. The data structure that we now observe is $O = (W, \Delta A, Y)$.

### Simulate Data

```{r}
# simulate simple data for tmle-shift sketch
n_obs <- 1000  # sample size
n_w <- 1  # just one baseline covariate for this example
tx_mult <- 2  # multiplier for the effect of W = 1 on the treatment

# baseline covariate -- simple, binary
W <- as.numeric(replicate(n_w, rbinom(n_obs, 1, 0.5)))

# set and organize treatment based on baseline W
A <- as.numeric(rnorm(n_obs, mean = tx_mult * W, sd = 1))

# create outcome as linear function of A, W + white noise
Y <- A + W + rnorm(n_obs, mean = 0, sd = 1)

# censoring based on covariates
C <- rbinom(n_obs, 1, plogis(W))

# treatment shift parameter
delta <- 0.5
```

---

## Methodology

__Note:__ In all examples below, the argument `eif_reg_type` is set to `"glm"`
(whereas the default is `"hal"`) when invoking an IPCW-TMLE. This is done in an
effort to speed along the computation process, as this argument controls the
manner in which a nuisance regression associated with the efficient influence
function of the IPCW-TMLE is fit. As stated in the documentation, setting this
argument to `"glm"` fits this EIF nuisance regression with a generalized linear
model (GLM; an assumption-laden parametric form) while the default value of
`"hal"` fits the nuisance regression with the highly adaptive lasso (HAL; a
recently developed nonparametric estimator) via the [`hal9001` R
package](https://github.com/tlverse/hal9001). In practice, we recommend
leaving this argument to the default and fitting the EIF nuisance regression
with HAL; however, this is significantly slower than simple using a GLM.

### Inverse Probability Weighting with Targeted Maximum Likelihood Estimation

```{r ipcw_tmle_ghal_Qglm_1}
tmle_hal_shift_1 <- txshift(W = W, A = A, Y = Y,
                            C = C, V = c("W", "Y"),
                            delta = delta,
                            fluc_method = "standard",
                            max_iter = 2,
                            ipcw_fit_args = list(fit_type = "glm"),
                            g_fit_args = list(fit_type = "hal",
                                              n_bins = 5,
                                              grid_type = "equal_mass",
                                              lambda_seq =
                                                exp(seq(-1, -9,
                                                        length = 300))),
                            Q_fit_args = list(fit_type = "glm",
                                              glm_formula = "Y ~ ."),
                            eif_reg_type = "glm"
                           )
summary(tmle_hal_shift_1)
```

When computing any such TML estimator, we may, of course, vary the regressions
used in fitting the nuisance parameters; however, an even simpler variation is
to fit the step for the fluctuation submodels with a _weighted_ method, simply
weighting each observation by an auxiliary covariate (often denoted $H_n$, and
sometimes called the "clever covariate" in the literature) rather than using
such a covariate directly in the submodel regression fit.

```{r ipcw_tmle_ghal_Qglm_2}
tmle_hal_shift_2 <- txshift(W = W, A = A, Y = Y,
                            C = C, V = c("W", "Y"),
                            delta = delta,
                            fluc_method = "weighted",
                            max_iter = 2,
                            ipcw_fit_args = list(fit_type = "glm"),
                            g_fit_args = list(fit_type = "hal",
                                              n_bins = 5,
                                              grid_type = "equal_mass",
                                              lambda_seq =
                                                exp(seq(-1, -9,
                                                        length = 300))),
                            Q_fit_args = list(fit_type = "glm",
                                              glm_formula = "Y ~ ."),
                            eif_reg_type = "glm"
                           )
summary(tmle_hal_shift_2)
```

### Interlude: Constructing Optimal Stacked Regressions with `sl3`

To easily incorporate ensemble machine learning into the estimation procedure,
we rely on the facilities provided in the [`sl3` R
package](https://sl3.tlverse.org). For a complete guide on using the `sl3` R
package, consider consulting https://sl3.tlverse.org, or https://tlverse.org for
the `tlverse` ecosystem, of which `sl3` is a major part.

```{r make_sl}
# SL learners to be used for most fits (e.g., IPCW, outcome regression)
lrnr_lib <- make_learner_stack("Lrnr_mean", "Lrnr_glm_fast", "Lrnr_xgboost")
sl_lrn <- Lrnr_sl$new(learners = lrnr_lib, metalearner = Lrnr_nnls$new())

# SL learners for conditional densities to be used for the propensity score fit
lrnr_dens_lib <- make_learner_stack(list("Lrnr_haldensify", n_bins = 5,
                                          grid_type = "equal_range",
                                          lambda_seq = exp(seq(-1, -7,
                                                               length = 100))),
                                    list("Lrnr_haldensify", n_bins = 5,
                                          bin_method = "equal_mass",
                                          lambda_seq = exp(seq(-1, -7,
                                                               length = 100))),
                                    list("Lrnr_haldensify", nbins = 10,
                                          bin_method = "equal_mass",
                                          lambda_seq = exp(seq(-1, -9,
                                                               length = 300)))
                                   )
sl_lrn_dens <- Lrnr_sl$new(learners = lrnr_dens_lib,
                           metalearner = Lrnr_solnp_density$new())
```

### Estimating the IPCW-TMLE with Optimal Stacked Regressions

Using the framework provided by the [`sl3` package](https://sl3.tlverse.org),
the nuisance parameters of the TML estimator may be fit with ensemble learning,
using the cross-validation framework of the Super Learner algorithm of
@vdl2007super. In principal, it would be desirable to estimate the parameter
using data-adaptive stacked regressions for the samplign mechanism (via
`ipcw_fit_args`, the treatment mechanism (via `g_fit_args`), and the outcome
mechanism (via `Q_fit_args`). This could be done with a call to the `txshift`
function like the following; however, we avoid running the following code chunk
for the sake of saving time in this vignette.


```{r ipcw_tmle_sl_1_not_run, eval=FALSE}
tmle_sl_shift_1 <- txshift(W = W, A = A, Y = Y,
                           C = C, V = c("W", "Y"),
                           delta = delta,
                           fluc_method = "standard",
                           max_iter = 2,
                           ipcw_fit_args = list(fit_type = "sl",
                                                sl_lrnrs = sl_lrn),
                           g_fit_args = list(fit_type = "sl",
                                             sl_lrnrs = sl_lrn_dens),
                           Q_fit_args = list(fit_type = "sl",
                                             sl_lrnrs = sl_lrn),
                           eif_reg_type = "glm"
                          )
summary(tmle_sl_shift_1)
```

As before, we may vary the regression for the submodel fluctuation procedure by
weighting each observation by the value of the auxiliary covariate rather than
using such an auxiliary covariate directly in the regression procedure. Note
that we again avoid running the following code chunk for the sake of saving time
in this vignette.

```{r ipcw_tmle_sl_2_not_run, eval=FALSE}
tmle_sl_shift_2 <- txshift(W = W, A = A, Y = Y,
                           C = C, V = c("W", "Y"),
                           delta = delta,
                           fluc_method = "weighted",
                           max_iter = 2,
                           ipcw_fit_args = list(fit_type = "sl",
                                                sl_lrnrs = sl_lrn),
                           g_fit_args = list(fit_type = "hal",
                                             n_bins = 5,
                                             grid_type = "equal_mass",
                                             lambda_seq =
                                               exp(seq(-1, -9,
                                                       length = 300))),
                           Q_fit_args = list(fit_type = "sl",
                                             sl_lrnrs = sl_lrn),
                           eif_reg_type = "glm"
                          )
summary(tmle_sl_shift_2)
```

### Estimating _Inefficient_ IPCW-TMLEs with Optimal Stacked Regressions

Utilizing the `sl3` R package in exactly the same manner, it is also possible to
estimate _inefficient_ IPCW-TML estimators with `txshift`. The inefficient
IPCW-TMLE performs nearly all of the same computations as the efficient version
of the estimator, eschewing only the iterative targeting procedure performed
over a more complex efficient influence function; the interested reader is
referred to @rose2011targeted2sd for further details on the differences between
the efficient and inefficient IPCW-TMLEs. By default, an efficient IPCW-TMLE is
computed when censoring is detected; however, this may be disable by setting
the `ipcw_efficiency` argument to `FALSE` like so

```{r ipcw_tmle_ghal_sl_inefficient}
tmle_sl_ineffic <- txshift(W = W, A = A, Y = Y,
                           C = C, V = c("W", "Y"),
                           delta = delta,
                           fluc_method = "standard",
                           ipcw_fit_args = list(fit_type = "sl",
                                                sl_lrnrs = sl_lrn),
                           g_fit_args = list(fit_type = "hal",
                                             n_bins = 5,
                                             grid_type = "equal_mass",
                                             lambda_seq =
                                               exp(seq(-1, -9,
                                                       length = 300))),
                           Q_fit_args = list(fit_type = "sl",
                                             sl_lrnrs = sl_lrn),
                           ipcw_efficiency = FALSE
                          )
summary(tmle_sl_ineffic)
```

Please note that in such cases, the value of `n_iter` in the resulting output
should be **_exactly zero_**, as this indicates that the iterative procedure
that is used to achieve efficiency has been skipped.

### Statistical Inference for Targeted Maximum Likelihood Estimates

For a discussion of the procedure for obtaining statistical inference for TML
estimators, the interested reader is referred to the introductory vignette of
this package. Here, we focus on addressing the issue of how censoring impacts
the inferential procedure...

```{r}
(ci_shift <- confint(tmle_hal_shift_1))
```

---

## _Advanced Usage:_ User-Specified Regressions

In some special cases it may be useful for the experienced user to compute the
censoring mechanism, treatment mechanism, and outcome mechanism regressions
separately (i.e., outside of the `txshift` wrapper function), instead
applying this user-facing wrapper only to invoke the _targeting_ steps involved
in computing the TML estimator for the treatment shift parameter. In such cases,
the optional arguments `ipcw_fit_spec`, `gn_fit_spec` and `Qn_fit_spec` may be
utilized. We present a case of using these arguments here:

```{r manual_ipcw_tmle}
# compute the censoring mechanism and produce IPC weights externally
pi_mech <- plogis(W)
ipc_weights_out <- (as.numeric(C == 1) / pi_mech)[C == 1]
ipcw_out <- list(pi_mech = pi_mech, ipc_weights = ipc_weights_out)

# compute treatment mechanism (propensity score) externally
## first, produce the down-shifted treatment data
gn_downshift <- dnorm(A - delta, mean = tx_mult * W, sd = 1)
## next, initialize and produce the up-shifted treatment data
gn_upshift <- dnorm(A + delta, mean = tx_mult * W, sd = 1)
## now, initialize and produce the up-up-shifted (2 * delta) treatment data
gn_upupshift <- dnorm(A + 2 * delta, mean = tx_mult * W, sd = 1)
## then, initialize and produce the un-shifted treatment data
gn_noshift <- dnorm(A, mean = tx_mult * W, sd = 1)
## finally, put it all together into an object like what's produced internally
gn_out <- as.data.table(cbind(gn_downshift, gn_noshift, gn_upshift,
                              gn_upupshift))[C == 1, ]
setnames(gn_out, c("downshift", "noshift", "upshift", "upupshift"))

# compute outcome regression externally
Qn_noshift <- (W + A - min(Y)) / diff(range(Y))
Qn_upshift <- (W + A + delta - min(Y)) / diff(range(Y))
Qn_noshift[Qn_noshift < 0] <- .Machine$double.neg.eps
Qn_noshift[Qn_noshift > 1] <- 1 - .Machine$double.neg.eps
Qn_upshift[Qn_upshift < 0] <- .Machine$double.neg.eps
Qn_upshift[Qn_upshift > 1] <- 1 - .Machine$double.neg.eps
Qn_out <- as.data.table(cbind(Qn_noshift, Qn_upshift))[C == 1, ]
setnames(Qn_out, c("noshift", "upshift"))

# invoke the wrapper function only to apply the targeting step
tmle_shift_spec <- txshift(W = W, A = A, Y = Y, delta = delta,
                           C = C, V = c("W", "Y"),
                           fluc_method = "standard",
                           max_iter = 2,
                           ipcw_fit_args = list(fit_type = "fit_spec"),
                           g_fit_args = list(fit_type = "fit_spec"),
                           Q_fit_args = list(fit_type = "fit_spec"),
                           ipcw_fit_spec = ipcw_out,
                           gn_fit_spec = gn_out,
                           Qn_fit_spec = Qn_out,
                           eif_reg_type = "glm")
summary(tmle_shift_spec)
```

---

## References

